---
title: "Lab 2"
date: January 30, 2022
author: Amy, Ksenia, and Rebecca
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)

#install.packages("fivethirtyeight")
#remotes::install_github("fivethirtyeightdata/fivethirtyeightdata")
#install.packages("waffle")
#install.packages("shadowtext")

library(fivethirtyeight)
library(fivethirtyeightdata)
library(tidyverse)
library(lubridate)
library(waffle)
library(shadowtext)

```

### 1.Start with the google trends data.
#### Visualize the change in trends for each hurricane over time in one plot with three scales. Map color to a discrete scale.

```{r google-trends}
trends <- google_trends

trends$date <- ymd(trends$date)

trends_pivot <- trends %>% 
	pivot_longer(
		cols = starts_with("hurricane"),
		names_to = c("Hurricane", "name"),
		values_to = "n",
		names_sep = "_",
		values_drop_na = TRUE,
		names_repair = "check_unique",
		names_prefix = "hurricane_"
	) 

ggplot(trends_pivot,aes(date, n)) +
	geom_line(aes(color = Hurricane), alpha = .5, size = 2) +
	scale_color_brewer(palette = "Dark2") +
	theme_minimal() 

ggplot(trends_pivot,aes(date, n)) +
	geom_line(aes(color = Hurricane), alpha = .5, size = 2) +
	scale_color_brewer(palette = "Dark1") +
	theme_minimal() 

ggplot(trends_pivot,aes(date, n)) +	
	geom_line(aes(color = Hurricane), alpha = .5, size = 2) +
	scale_color_discrete() +
	theme_minimal()
```

#### Visualize the trends again, this time with solid shapes (if your first plot was already solid shapes, use lines). Hint, geom_area can be difficult to work with if you have a grouping factor. Try geom_ribbon instead.
```{r google-trends2}
ggplot(trends_pivot, aes(x = date, y = n)) +
	geom_ribbon(aes(x = date, y = n, ymin = n - 1, ymax = n + 1, fill = Hurricane, color = Hurricane, linetype = Hurricane)) +
	geom_area(aes(fill = Hurricane)) +
	theme_minimal()
```

#### Visualize the same relation, but change the scales so color is mapped to a continuous scale (the other scales should be discrete).

```{r google-trends3}
ggplot(trends_pivot, aes(x = date, y = Hurricane, fill = n)) +
	geom_tile() +
	scale_fill_continuous(type = "viridis")
```

#### Create a data frame (using tibble() or data.frame()) that includes the dates that each hurricane made landfall. Annotate one of your first two plots to include these reference points with vertical lines (Hint: use geom_vline(), specifying the data source for this layer as this new dataset). You do not have to include text annotations - just the vertical lines, but if you would like to try, look into geom_text() or geom_label(). 
```{r google-trends4}

# Didn't include Jose b/c no reference line is needed since it didn't make landfall

landfall <- tibble(Hurricane = c("Harvey", 
                                  "Irma", 
                                  "Maria"), 
                    Date = c("August 25, 2017", 
                             "September 10, 2017", 
                             "September 20, 2017"), 
                    numdate = c("2017-08-25", 
                                "2017-09-10", 
                                "2017-09-20"))

# Overlaid vertical reference lines on top of the filled line graph 

ggplot(trends_pivot, 
       aes(x = date, 
           y = n)) +
	geom_ribbon(aes(x = date, 
	                y = n, 
	                ymin = n - 1, 
	                ymax = n + 1, 
	                fill = Hurricane, 
	                color = Hurricane, 
	                linetype = Hurricane)) +
	geom_area(aes(fill = Hurricane)) +
	theme_minimal() +
  geom_vline(aes(xintercept = ymd(numdate)), 
             data = landfall, 
             lty = 2,
             alpha = 0.4)  

# Added labels to vertical reference lines
ggplot(trends_pivot, 
       aes(x = date, 
           y = n)) +
	geom_ribbon(aes(x = date, 
	                y = n, 
	                ymin = n - 1, 
	                ymax = n + 1, 
	                fill = Hurricane, 
	                color = Hurricane, 
	                linetype = Hurricane)) +
	geom_area(aes(fill = Hurricane)) +
	theme_minimal() +
  scale_y_continuous(expand = c(0,0)) + 
  labs(x = "Date", 
       y = "Google Search Interest") +
  geom_vline(aes(xintercept = ymd(numdate)), 
             data = landfall, 
             lty = 2,
             alpha = 0.4) +
  geom_label(data = landfall, 
             aes(x = ymd(numdate), 
                 y = 75, 
                 label = c("Harvey
 Landfall 
 Aug. 25", 
                           "Irma 
 Landfall
 Sep. 10", 
                           "Maria 
 Landfall
 Sep. 20")), 
           color = "gray50",
           size = 2.5)
```



```{r tv_states}
tv_states$date <- ymd(tv_states$date)

states_pivot <- tv_states %>% 
	pivot_longer(!date,
		names_to = "state",
		values_to = "share")

states_pivot <- states_pivot %>% 
  mutate(
    state = case_when(
      state == "florida" ~ "Florida",
      state == "texas" ~ "Texas",
      state == "puerto_rico" ~ "Puerto Rico"
    ))

states_pivot$state <- factor(states_pivot$state, levels = c("Florida", "Texas", "Puerto Rico"))

labels_part2 <- tibble(event = c("Harvey", 
                                  "Irma", 
                                  "Maria",
                                 "Las Vegas"), 
                    Date = c("August 25, 2017", 
                             "September 10, 2017", 
                             "September 20, 2017",
                             "October 1, 2017"), 
                    numdate = c("2017-08-25", 
                                "2017-09-10", 
                                "2017-09-20",
                                "2017-10-01"))

labels3 <- tibble(x = c(ymd("2017-08-29"), 
                        ymd("2017-09-09"), 
                        ymd("2017-09-30")), 
                 y = c(0.9, 1.6, 1.6),
                 label = c("Texas", 
                           "Florida", 
                           "Puerto Rico"))

ggplot(states_pivot, 
       aes(x = date, 
           y = share)) +
	geom_ribbon(aes(x = date, 
	                y = share, 
	                xmin = ymd("2017-08-15"), 
	                ymin = share, 
	                ymax = share,
	                fill = state, 
	                color = state, 
	                linetype = state)) +
	geom_area(aes(fill = state), color = "white", size = 0.4) +
	theme_minimal() +
  scale_fill_manual(values = c("#d23586",  "#de743c", "#10c8dd")) +
  scale_y_continuous(expand = c(0,0), labels = function(share) paste0(share,"%")) + 
  labs(x = element_blank(), 
       y = "Share of sentences", 
       title = "National cable news network", 
       caption = "Includes Bloomberg, CNBC, CNN, Fox Business, Fox News and MSNBC.") +
  geom_vline(aes(xintercept = ymd(numdate)), 
             data = labels_part2, 
             linetype = 'dotted',
             alpha = 0.4) +
  geom_label(data = labels_part2, 
             aes(x = c(ymd(numdate[1]), 
                       ymd("2017-09-14"), 
                       ymd(numdate[3]), 
                       ymd(numdate[4])), 
                 y = c(3.2, 3.4, 2.4, 3.05),
                 label = c("Harvey
 landfall", 
                           "Irma 
 landfall", 
                           "Maria 
 landfall",
                            "Las Vegas
 shooting"),
 fill = NULL), 
                 color = "gray50",
                 size = 4,
                 fontface = "italic",
                 label.size = NA) +
  geom_shadowtext(data = labels3, 
                  aes(x = x, 
                      y = y,
                      label = label), 
                  color = "black",
                  bg.color ="white",
                  size = 5) +
  theme(panel.grid.minor.x = element_blank(), 
        panel.grid.minor.y = element_blank(),
        axis.line.x = element_line(size = 0.5, color = "grey40"),
        plot.title = element_text(hjust = 0.5, face = "bold"), 
        axis.title.y = element_text(face = "bold"), 
        axis.text = element_text(color = "gray50", size = 10),
        plot.caption = element_text(hjust = 0),
        legend.position = "none")
```

### 3. Use the comic_characters dataset to visualize the proportion of different genders using at least three different methods. With your final visualization, make any additional modifications you think would help reduce the cognitive load necessary to comprehend the figure.

```{r comic-prep}
# Load the data
comic <- comic_characters

# Take a look at the ns
comic %>% 
  group_by(sex) %>% 
  summarize(n = n())

# Prelim plot
comic %>% 
  ggplot(aes(sex)) +
  geom_bar() +
  coord_flip()

# New tibble w/ proportions by gender
genderprop <- comic %>% 
  group_by(sex) %>% 
  summarize(n = n(), 
            proportion = (n()/nrow(comic)))

# Set gender as a factor
genderprop <- genderprop %>% 
  mutate(sex = as_factor(sex)) 

# Change "NA" Gender to "Missing"
genderprop$sex <- fct_explicit_na(genderprop$sex, na_level = "Missing")

# Remove "Characters" from levels of gender variable
genderprop$sex <- recode_factor(genderprop$sex, 
                                `Male Characters` = "Male", 
                                `Female Characters` = "Female", 
                                `Agender Characters` = "Agender",
                                `Genderless Characters` = "Genderless",
                                `Genderfluid Characters` = "Genderfluid",
                                `Transgender Characters` = "Transgender")
```

#### Plot 1 - Vertical bar graph
```{r comic-plot1}

genderprop %>% 
  ggplot(aes(fct_reorder(sex, proportion), proportion)) +
  geom_col(fill = "#DA6541",
           color = "#FFFFFF") +
  coord_flip() + 
  theme_minimal()
```

#### Plot 2 - Waffle chart
```{r comic-plot2}
# Waffle Chart v1. Challenging due to percentages and rounding. If 100 squares are used, none can be assigned to Agender, Genderless, Genderfluid, or Transgender. In this first version, I still included these categories in the key. However, that could be confusing to the reader since colors are assigned to the categories but they aren't in the graph. The default colors (set2) are used. 
genderpropwaffle1 <- c(`Male = 71%` = 71, 
                      `Female = 25%` = 25, 
                      `Missing = 4%` = 4, 
                      `Agender = <1%` = 0, 
                      `Genderfluid = <1%` = 0, 
                      `Genderless = <1%` = 0, 
                      `Transgender = <1%` = 0)

waffle(genderpropwaffle1)

# Waffle v2. Top three categories. I don't like this one because not all gender categories are represented.

genderpropwaffle2 <- c(`Male = 71%` = 71, 
                       `Female = 25%` = 25, 
                       `Missing = 4%` = 4)

waffle(genderpropwaffle2, 
       colors = c("#4B9EA8", 
                  "#DA6541",
                  "#5F666E"),
       xlab = "1 sq = 1%")

# Waffle v3. Better but still not super happy with it. Would need to use a lot of waffle squares in order for all categories to be represented visually. I think this could work but would need further notes for the reader. I also was interested in seeing if I could a get a glyph to work (specifically the mask one) but did not have luck.

#library(extrafont)
#font_import()
#loadfonts()
#fonts()[grep("Awesome", fonts())]

genderpropwaffle3 <- c(`Male = 71%` = 71, 
                       `Female = 25%` = 25, 
                       `Missing = 4%` = 4, 
                       `Agender = <1%` = 0, 
                       `Genderfluid = <1%` = 0, 
                       `Genderless = <1%` = 0, 
                       `Transgender = <1%` = 0)

waffle(genderpropwaffle3, 
       colors = c("#4B9EA8", 
                  "#DA6541",
                  "#5F666E", 
                  "#FFFFFF", 
                  "#FFFFFF", 
                  "#FFFFFF", 
                  "#FFFFFF"))
```

#### Plot 3 - Points to indicate percentage instead of bars 
```{r comic-plot3}
# I prefer this over the bar graph because the points for the gender categories with percentages <1% are visible. I used a minimal theme, one color, and clear labeling to reduce cognitive load.

unique(comic$year)
unique(comic$publisher)

genderprop %>% 
  ggplot(aes(fct_reorder(sex, proportion), proportion)) +
  geom_point(color = "#DA6541") +
  coord_flip() +
  theme_minimal() + 
  labs(title = "Gender Representation in DC and Marvel Comics",
       caption = "Based on character data from 1939 to 2013.",
       x = "Gender",
       y = "Percentage") +
  theme(plot.title.position = "plot") + 
  scale_y_continuous(labels = scales::percent)
```

